---
title: "Assignment 3, Tasks 3 & 4, ESM-206"
author: "Sean Denny"
date: "11/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Functions used:

here()
clean_names()
geom_histogram()
geom_boxplot()
geom_col()
filter()
mutate()
separate()
case_when()
count() (combines `group_by()` and summarize n() all in one function).
kable()


## Assignment:

Load packages and data, and convert column names to lower snake case:

```{r Load packages and data, message = FALSE}

library(here)
library(janitor)
library(kableExtra)
library(lubridate)
library(RColorBrewer)
library(tidyverse)

# Is this a problem, because some of the functions might conflict with each other? Are conflicting functions common across packages?

urchins <- read_csv(here("data","urchins.csv")) %>%
  # convert the names to lower snake case
  clean_names()
```

### Task 3

#### Section 1

Explore the data:

```{r}

# Exploratory histogram of size by urchin species (red vs. purple), faceted by common name:

urchins_histo <- ggplot(data = urchins, aes(x = size)) +
  geom_histogram(bins = 25) +
  facet_wrap(~common_name) +
  theme_minimal() +
  labs(x = "urchin diameter (cm)")

urchins_histo

# Exploratory box plot with common name on x-axis and size on y-axis:

urchins_box <- ggplot(data = urchins, aes(x = common_name, y = size)) +
  geom_boxplot() +
  labs(x = " ", y = "urchin diameter (cm)") +
  theme_minimal()

urchins_box

```

Descriptive statistics table:

```{r}

urchin_table <- urchins %>% 
  group_by(common_name) %>% 
  summarize(mean_size = mean(size), median_size = median(size), standard_deviation = sd(size), sample_size = n()) %>% 
  kable(col.names = c("species", "mean size (cm)", "median size (cm)", "standard deviation (cm)", "sample size")) %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

# bootstrap_options = "striped" gives it the alternativing gray and white backdrop across rows. Could just leave it out to keep things simpler. In this case, this option is probably unnecssary.

urchin_table

```

#### Section 2

Explore purple urchin counts by site and year:

```{r}

purple_urchins <- urchins %>% 
  filter(common_name == "Purple Urchin") %>% 
  mutate(date = mdy(date)) %>% 
  separate(col = date,
           into = c("year", "month", "day"),
           sep = "-",
           remove = FALSE) %>%
mutate(site_full = case_when(
  site == "NAPL" ~ "Naples Reef",
  site == "AQUE" ~ "Arroyo Quemado Reef",
  site == "IVEE" ~ "Isla Vista Reef",
  site == "CARP" ~ "Carpinteria Reef",
  site == "MOHK" ~ "Mohawk Reef",
))
```

Make a summary table showing urchin count by site and year

```{r}
# So I think I want to have three columns, site, year, count, with a set of rows for each site covering all the years.

# counts by site and year

purple_urchins_site_year <- purple_urchins %>% 
  count(site_full, year) %>% 
  rename("sample_size" = n)
  
```

Make an exploratory graph of counts of purple urchins recorded at chea site over time for years in the study.

```{r}

urchins_graph <- ggplot(data = purple_urchins_site_year, aes(x = year, y = sample_size)) +
  geom_col() +
  facet_wrap(~site_full) +
  theme_minimal() +
   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(y = "count", x = "", title = "Purple urchin counts by year and site") +
  theme(plot.title = element_text(hjust = 0.5))
  
urchins_graph

```

### Task 4














