---
title: "A3_T4_SD_2020"
author: "Sean Denny"
date: "11/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Functions used:

here()
clean_names()
geom_histogram()
geom_boxplot()
geom_col()
filter()
mutate()
separate()
case_when()

## Assignment:

Load packages and data, and convert column names to lower snake case:

```{r Load packages and data, message = FALSE}

library(here)
library(janitor)
library(kableExtra)
library(lubridate)
library(RColorBrewer)
library(tidyverse)

# Is this a problem, because some of the functions might conflict with each other? Are conflicting functions common across packages?

urchins <- read_csv(here("data","urchins.csv")) %>%
  # convert the names to lower snake case
  clean_names()
```

Compare mean size of purple urchins in control and annual sections of Naple Reef in Janary 2010 using a two-sample t-test.

Create a dataset that only has info for Naples Reef for 2010.

```{r}

naples_2010 <- urchins %>% 
  filter(common_name == "Purple Urchin") %>% 
  mutate(date = mdy(date)) %>% 
  separate(col = date,
           into = c("year", "month", "day"),
           sep = "-",
           remove = FALSE) %>%
mutate(site_full = case_when(
  site == "NAPL" ~ "Naples Reef",
  site == "AQUE" ~ "Arroyo Quemado Reef",
  site == "IVEE" ~ "Isla Vista Reef",
  site == "CARP" ~ "Carpinteria Reef",
  site == "MOHK" ~ "Mohawk Reef",
)) %>% 
  filter(site_full == "Naples Reef" & year == "2010") %>% 
  select(size, treatment)
```

Explore the data using histograms and q-q plots, per treatment:

Histograms:

```{r}
naples_histogram <-ggplot(data = naples_2010, aes(x = size)) +
  geom_histogram(fill = "purple", color = "black", bins = 10) +
  facet_wrap(~treatment)

naples_histogram
```

Q-q plots:

```{r}
naples_qq <- ggplot(data = naples_2010, aes(sample = size))+
  geom_qq() +
  facet_wrap(~ treatment)

naples_qq
```

Create a summary table of descriptive statistics for annual and control groups.

```{r}

naples_table <- naples_2010 %>% 
  group_by(treatment) %>% 
  summarize(mean_size = mean(size), standard_dev = sd(size), sample_size = n()) %>% 
  mutate(treatment = case_when(
    treatment == "ANNUAL" ~ "annual",
    treatment == "CONTROL" ~ "control"
  )) %>% 
  kable(col.names = c("treatment type", "mean size (cm)", "standard deviation (cm)", "sample size")) %>% 
    kable_styling(bootstrap_options = "striped", full_width = FALSE) # Could just erase bootstrap_option = "striped" to remove the gray alternating stripe.

naples_table

```

Perform a two-sample t-text to compare means between the two groups

```{r}

# First create subsets of data for each group:

naples_control <- naples_2010 %>% 
  filter(treatment == "CONTROL") %>% 
  select(size)

naples_annual <- naples_2010 %>% 
  filter(treatment == "ANNUAL") %>% 
  select(size)

t.test(x = naples_control, y = naples_annual)


```
